Reconstructing the last known movements of one of Nature's giants

### 2018-05-02 manuscript/CurrentBiology/blue-whale-ms.tex
Edit abstract

\section{Summary}
Understanding animal movements is crucial for effective conservation, especially as species ranges shift due to climate change. 
Animal tissues housed in museum collections provide biochemical records of behaviour that can be used to reconstruct individual-level movements over long timescales and under historical climatic and ecological conditions, but interpreting these records can be challenging. 
Here we combine stable isotope analyses of baleen with agent-based simulation models, to reconstruct seven years of movements of an iconic female blue whale (\textit{Balaenoptera musculus}) that stranded off Ireland in 1891 and is now on display at the Natural History Museum, London. 
We provide the longest continuous record of movement for an individual blue whale in the north Atlantic. 
We show that the whale spent at least two years resident in sub-tropical waters as a juvenile, before commencing annual latitudinal migrations consistent with patterns of movements of blue whales in the northeast Atlantic inferred from observations, tagging and historic whaling data. 
The isotopic record is consistent with observations of opportunistic feeding throughout the year in blue whales, and suggests pregnancy and calf-rearing in the year before death.
Historic tissue collections can provide unique insight into individual-level movement behaviours associated with changing climate, ecological conditions and anthropogenic pressures, and reveal aspects of life history. 
Combining biochemical tracers with simulation modeling unlocks the information recorded in these irreplaceable archives.

>>>>

\section{Summary}
The spatial ecology of rare, migratory, oceanic mammals is challenging to study directly, leading to distinct knowledge gaps associated with multi-year migratory behavior at the individual scale.
Animal tissues housed in museum collections provide biochemical records of behavior that can be used to reconstruct individual-level movements over long timescales and under historical climatic and ecological conditions, but interpreting these records is difficult. 
Many baleen whales face severe conservation threats, therefore knowledge of their movements and space use over multiple years are needed. 
Here we combine stable isotope analyses of baleen with novel agent-based simulation models, to infer seven years of movements of an iconic blue whale (\textit{Balaenoptera musculus}) that stranded off Ireland in 1891 and is now on display at the Natural History Museum, London. 
We identify three distinct behavioral states that we interpret as residency in sub-tropical waters (potentially as a juvenile), three annual cycles of seasonal latitudinal migrations and pregnancy and calf-rearing in the year before death.
Our approach combining simulation modeling and stable isotope data yields individual-level movement inferences that are consistent with the current population-based understanding of the movement ecology of blue whales in the northeast Atlantic ocean. 
Combining biochemical tracers with simulation modeling unlocks the information recorded in historic collections, providing a link between traditional population-based assessments of animal distributions and more recent methodologies drawing inferences from high-resolution observations of relatively few individuals. 
This approach can be adapted to infer movements in any animal where underlying stable isotope models and incrementally grown tissues are available.
